---
title: "FiberMargin benchmarks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FiberMargin benchmarks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(fibermargin)
```

## Evaluation contract

FiberMargin is evaluated as a coordinate-and-label repair operator.
Each method receives:

- coordinates (`xy`),
- one observed categorical field (`labels`),
- optional specimen identifiers (`samples`),
- optional evaluation masks (`boundary`, `sparse`, `regions`).

Reference labels are only used after refinement for scoring.
Primary metrics are:

- `accuracy` and `ari`,
- `correction_recall` (recovered errors),
- `damage_rate` (introduced errors),
- `worst_recall` (min class recall),
- `boundary_accuracy` and `sparse_region_accuracy`.

```{r real-summary}
available_spatial_benchmarks()
```

For bundled benchmarks, this package stores corrupted scenarios for DLPFC and CRC
with reproducible recipes and metadata in object-level `design` tables.
`load_spatial_benchmark()` returns a `spatial_refinement_benchmark`, ready for use
with `benchmark_spatial_refiners()`.

```{r scenario-sanity, eval=FALSE}
dlpfc <- load_spatial_benchmark("dlpfc", scenario = 1L)
crc <- load_spatial_benchmark("crc", scenario = "CRC_random_25_r1")
c(dlpfc_n = nrow(dlpfc$xy), dlpfc_sections = nlevels(dlpfc$samples))
c(crc_n = nrow(crc$xy), crc_classes = nlevels(crc$truth))
```

## One benchmark run

```{r simple-benchmark, eval=FALSE}
sim <- simulate_spatial_domains(
  n = 50000L, pattern = "jagged_stripes", noise_type = "boundary",
  density_profile = "extreme", seed = 7L
)

results <- benchmark_spatial_refiners(
  sim,
  methods = list(
    FiberMargin = refine_spatial_labels
  ),
  include_initial = TRUE
)
```

```{r quick-result, eval=FALSE}
results
```

Run this block to inspect the benchmark output generated on your machine.
For full protocol numbers, run the scripts in `benchmarks/` that generate the
comparison tables used in the manuscript.

## Stress-test geometries

```{r simulation-catalogue, eval=FALSE}
jagged <- simulate_spatial_domains(
  n = 50000L, pattern = "jagged_stripes", noise_type = "boundary",
  density_profile = "extreme", seed = 1L
)
gradient <- simulate_gradient_regions(
  n = 50000L, minority = 0.05, dimensions = 3, samples = 3L, seed = 2L
)
tubes <- simulate_complex_spatial_domains(
  n = 50000L, shape = "tubular_network", density_profile = "extreme", seed = 3L
)
volume <- simulate_volumetric_domains(
  n = 50000L, shape = "helical_channels", acquisition = "irregular_z", seed = 4L
)
```

`benchmark_spatial_refiners()` is suitable for direct, apples-to-apples
comparisons across these geometries because all returned masks and diagnostics are
scored against the same `truth`, `boundary`, and sparse-region masks.

```{r benchmark-api, eval=FALSE}
benchmark_spatial_refiners(jagged)
```

## Reproducible scripts

```sh
Rscript benchmarks/run_publication_benchmarks.R
Rscript benchmarks/run_fibermargin_publication_benchmarks.R
Rscript benchmarks/run_general_mask_repair_benchmarks.R
```

These scripts create full result matrices, summary tables, and plotting assets used in
the paper workflow.
