---
title: "FiberMargin benchmarks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FiberMargin benchmarks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(fibermargin)
```

## Evaluation contract

FiberMargin is evaluated as a coordinate-and-label repair operator. Each method
receives a coordinate matrix, one observed categorical field, and optional
specimen identifiers. Reference annotations are used only after repair to score
controlled corruptions. Accuracy and adjusted Rand index (ARI) are primary;
correction recall and damage describe how aggressively a method edits the input.

The predefined real-coordinate protocols contain 60 corruptions of a 19-class CRC
Visium HD annotation and 45 corruptions over five consecutive MERFISH planes.
Each protocol preserves at least one observed exemplar of every reference class.
The direct fixed controls are alpha-expansion Potts, a coordinate CRF, harmonic
propagation, multiscale mode filtering, Potts-like ICM, fixed SpaGCN and GraphST
correction rules, and an 8-nearest-neighbour modal filter.

| Dataset | Units | FiberMargin accuracy | FiberMargin ARI | Best accuracy | Best ARI |
|:--|--:|--:|--:|:--|:--|
| CRC | 60 | 0.8600 | 0.7812 | 0.8604 (Multiscale) | 0.7827 (Multiscale) |
| MERFISH | 45 | 0.8886 | 0.7951 | 0.8843 (Alpha-expansion Potts) | 0.7888 (Alpha-expansion Potts) |

The complete per-case ledgers, control configuration, stored-input identities,
and run manifests are in
`benchmarks/results/fibermargin_final_enclosure_real_matrix_v2/`. These are
controlled label-recovery results; repeated corruptions of one annotation are
not independent biological specimens.

## Report all outcomes

The exported evaluator reports primary accuracy and ARI together with boundary,
sparse-region, and worst-class performance, correction recall, and damage.

```{r evaluation, eval=FALSE}
sim <- simulate_spatial_domains(
  n = 50000,
  pattern = "jagged_stripes",
  noise_type = "boundary",
  density_profile = "extreme",
  seed = 7
)

refined <- refine_spatial_labels(sim$xy, sim$labels, sim$samples, workers = 4)
evaluate_spatial_refinement(
  truth = sim$truth,
  initial = sim$labels,
  refined = refined,
  boundary = sim$boundary,
  sparse = sim$sparse,
  method = "FiberMargin"
)
```

`damage_rate` is the fraction of initially correct labels made incorrect by
repair. It complements `correction_recall`, the fraction of initially incorrect
labels recovered. A method with a high correction recall can still reduce
accuracy if it damages too many correct labels.

## Stress-test geometries

The package exposes planar, gradient, complex-shape, and volumetric simulators.
They return coordinates, the observed labels, evaluation-only truth, specimen
membership, and boundary/sparse masks, so every method can be scored under the
same conditions.

```{r simulation-catalogue, eval=FALSE}
jagged <- simulate_spatial_domains(
  n = 50000, pattern = "jagged_stripes", noise_type = "boundary",
  density_profile = "extreme", seed = 1
)
gradient <- simulate_gradient_regions(
  n = 50000, minority = 0.05, dimensions = 3, samples = 3, seed = 2
)
tubes <- simulate_complex_spatial_domains(
  n = 50000, shape = "tubular_network", density_profile = "extreme", seed = 3
)
volume <- simulate_volumetric_domains(
  n = 50000, shape = "helical_channels", acquisition = "irregular_z", seed = 4
)
```

`benchmark_spatial_refiners()` applies named repair functions to one simulated
object with the same evaluation masks. Its default is FiberMargin.

```{r benchmark-api, eval=FALSE}
benchmark_spatial_refiners(jagged)
```

## Reproduce the fixed analyses

The scripts below reproduce the direct matrix and the focused mechanism and
component audits on the stored input fields.

```sh
Rscript benchmarks/reconcile_isolation_guard_real_matrix.R
Rscript benchmarks/audit_two_sided_enclosure_real.R
Rscript benchmarks/audit_fibermargin_component_reduction_real.R
```

Use `R_LIBS_USER` to point these commands to the tested local package library
when working from a source checkout.
