---
title: "FiberMargin"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fibermargin}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 3
)
library(fibermargin)
```

## What FiberMargin does

FiberMargin is a deterministic, training-free routine that repairs noisy categorical
labels using only spatial coordinates and observed class labels.

- `xy`: 2D/3D coordinates, one row per spot.
- `labels`: observed class labels.
- `samples`: optional section/specimen identifier. Different specimens are refined
  independently.
- `workers`: optional CPU budget.

The returned value is a factor with the same levels and row names as `labels`, and
attributes with pointwise diagnostics.

## Quick start

```{r quick-start}
sim <- simulate_gradient_regions(
  n = 4000L,
  minority = 0.10,
  samples = 2L,
  seed = 12L
)

refined <- refine_spatial_labels(
  xy = sim$xy,
  labels = sim$labels,
  samples = sim$samples,
  workers = 1L
)

c(
  input_accuracy = mean(sim$labels == sim$truth),
  refined_accuracy = mean(refined == sim$truth),
  changed = mean(refined != sim$labels)
)
```

## Diagnostics

```{r diagnostics}
str(
  list(
    candidate = head(attr(refined, "candidate")),
    changed = head(attr(refined, "changed")),
    repair_margin = head(attr(refined, "repair_margin"))
  )
)
```

- `candidate`: local best candidate class at each spot.
- `margin_score`: local support contrast for the candidate.
- `required`: adaptive acceptance threshold.
- `repair_margin`: `margin_score - required`; accepted changes have non-negative
  margin.
- `atlas_dispersion`: directional view disagreement.
- `isolation`: local isolation protection factor.
- `changed`: whether the output label differs from the input label.

## Clean a categorical mask

For 2D masks and 3D arrays, mask cleaning uses the same coordinates-only
refiner. Missing cells are preserved.

```{r clean-mask}
truth <- matrix(rep(c("background", "region"), each = 3200L), nrow = 80)
noisy <- corrupt_categorical_mask(
  truth,
  mechanism = "boundary",
  rate = 0.15,
  seed = 19L
)
cleaned <- clean_categorical_mask(noisy, workers = 1L)

evaluate_mask_cleaning(
  reference = truth,
  initial = noisy,
  cleaned = cleaned,
  method = "FiberMargin"
)
```

## Benchmark helper

```{r benchmark-example, eval=FALSE}
bench <- simulate_spatial_domains(
  n = 6000L,
  pattern = "jagged_stripes",
  noise = 0.20,
  samples = 2L,
  seed = 5L
)

benchmark_spatial_refiners(
  bench,
  methods = list(
    FiberMargin = refine_spatial_labels,
    InitialOnly = function(xy, labels, ...) labels
  ),
  include_initial = TRUE,
  seed = 1L
)
```

## Real-data flow

```{r real-data}
available_spatial_benchmarks()
dlpfc <- load_spatial_benchmark("dlpfc", scenario = 1L)

refined <- refine_spatial_labels(
  dlpfc$xy,
  dlpfc$labels,
  samples = dlpfc$samples,
  workers = 2L
)

evaluate_spatial_refinement(
  truth = dlpfc$truth,
  initial = dlpfc$labels,
  refined = refined,
  boundary = dlpfc$boundary,
  sparse = dlpfc$sparse,
  method = "FiberMargin"
)
```

## Practical notes

- Keep defaults unless you have a strong reason to modify internal options.
- Use `workers` when processing many independent specimens.
- Report both aggregate scores (`accuracy`, `ari`) and failure-mode scores
  (`damage_rate`, `correction_recall`, `boundary_accuracy`, `sparse_region_accuracy`).
