---
title: "Getting started with marginSVM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with marginSVM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 3)
library(marginSVM)
```

## Refine spatial labels

marginSVM needs spatial coordinates and an initial label for every location. A
tissue or section identifier is optional. The example below contains four areas
whose dominant classes are A, B, B, and C, with minority labels mixed into each
area.

```{r refine}
sim <- simulate_gradient_regions(
  n = 4000,
  minority = 0.25,
  samples = 2,
  density_profile = "strong",
  seed = 12
)

refined <- refine_spatial_svm(
  xy = sim$xy,
  labels = sim$labels,
  samples = sim$samples
)

c(
  input_accuracy = mean(sim$labels == sim$truth),
  refined_accuracy = mean(refined == sim$truth)
)
```

Tiles and model settings are automatic. Different tissues never share graph
neighbors or local models.

## Visualize the correction

```{r spatial-plot, echo=FALSE, fig.cap="Reference regions, noisy input labels, and marginSVM output for the first tissue."}
rows <- sim$samples == levels(sim$samples)[1]
palette <- c(A = "#2A9D8F", B = "#E9C46A", C = "#E76F51")
old_par <- par(mfrow = c(1, 3), mar = c(1.5, 1.5, 2, 0.5))
plot(sim$xy[rows, ], col = palette[sim$truth[rows]], pch = 16, cex = 0.35,
     axes = FALSE, asp = 1, main = "Reference")
plot(sim$xy[rows, ], col = palette[sim$labels[rows]], pch = 16, cex = 0.35,
     axes = FALSE, asp = 1, main = "Noisy labels")
plot(sim$xy[rows, ], col = palette[refined[rows]], pch = 16, cex = 0.35,
     axes = FALSE, asp = 1, main = "marginSVM")
par(old_par)
```

## Inspect diagnostics

The result is a factor with the same levels as the input. Per-location evidence
and execution information are attached without changing the ordinary vector
interface.

```{r diagnostics}
head(data.frame(
  label = refined,
  confidence = attr(refined, "confidence"),
  margin = attr(refined, "margin"),
  local_support = attr(refined, "local_support")
))

c(
  tiles = attr(refined, "tiles"),
  workers = attr(refined, "workers")
)
attr(refined, "abstained_samples")
```

Low margin indicates that the final decision lies near a competing class.
`local_support` measures agreement of the input label with nearby labels. These
scores are diagnostics rather than calibrated biological probabilities.

## Use measured spatial data

For a Seurat object, extract coordinates and pass the annotation vector directly.
The following code is not evaluated because Seurat is not a package dependency.

```{r seurat, eval=FALSE}
library(Seurat)

xy <- GetTissueCoordinates(object)[, c("x", "y")]
labels <- object$wsi_annotation_name
object$marginSVM <- refine_spatial_svm(xy, labels)
```

For multiple sections, concatenate coordinates and labels and pass a matching
section identifier:

```{r multiple-tissues, eval=FALSE}
refined <- refine_spatial_svm(
  xy = coordinates,
  labels = initial_labels,
  samples = section_id
)
```

## CPU and optional accelerators

The package always includes the multithreaded C++ CPU backend. Large datasets are
automatically divided into overlapping 2D or 3D tiles. Use `workers` only when a
specific CPU limit is needed.

```{r backends}
spatial_svm_backend_capabilities()
```

CUDA and Metal are provider interfaces. If no provider is registered, `"auto"`
uses CPU and an explicitly requested unavailable accelerator warns before falling
back to CPU.
