---
title: "FiberMargin workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FiberMargin workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 3
)
library(fibermargin)
```

## Refine spatial labels

FiberMargin refines observed categorical labels on spatial coordinates.
`xy` is a 2D/3D coordinate matrix and `labels` are the observed classes.
`samples` is optional and enables independent processing by specimen.

```{r refine}
sim <- simulate_gradient_regions(
  n = 2400L,
  minority = 0.10,
  samples = 2L,
  density_profile = "strong",
  seed = 12
)

refined <- refine_spatial_labels(
  xy = sim$xy,
  labels = sim$labels,
  samples = sim$samples,
  workers = 1L
)

c(
  input_accuracy = mean(sim$labels == sim$truth),
  refined_accuracy = mean(refined == sim$truth)
)
```

The method is deterministic and uses one fixed internal rule set. Increasing
`workers` enables parallel processing across specimens.

## Clean a categorical mask

For 2D masks and 3D arrays, missing cells are preserved in the output.

```{r clean-mask}
reference_mask <- matrix(rep(c("background", "region"), each = 800L), nrow = 40)
noisy_mask <- corrupt_categorical_mask(reference_mask, mechanism = "impulse", rate = 0.15, seed = 19)
cleaned_mask <- clean_categorical_mask(noisy_mask, workers = 1L)

evaluate_mask_cleaning(reference_mask, noisy_mask, cleaned_mask)[, c(
  "initial_accuracy", "accuracy", "mean_iou", "mean_boundary_iou",
  "correction_recall", "damage_rate"
)]
```

`accuracy_gain` follows the exact identity
`accuracy_gain = repaired_fraction - damaged_fraction`.
Reporting both repair and damage avoids over-interpreting large gains from very aggressive updates.

## Visualize the correction

```{r spatial-plot, echo=FALSE, fig.cap="Reference regions, noisy input labels, and FiberMargin output for one specimen."}
rows <- sim$samples == levels(sim$samples)[1L]
palette <- c(A = "#0072B2", B = "#E69F00", C = "#009E73")
old_par <- par(mfrow = c(1, 3), mar = c(1.5, 1.5, 2, 0.5))
plot(sim$xy[rows, ], col = palette[sim$truth[rows]], pch = 16, cex = 0.45,
     axes = FALSE, asp = 1, main = "Reference")
plot(sim$xy[rows, ], col = palette[sim$labels[rows]], pch = 16, cex = 0.45,
     axes = FALSE, asp = 1, main = "Noisy labels")
plot(sim$xy[rows, ], col = palette[refined[rows]], pch = 16, cex = 0.45,
     axes = FALSE, asp = 1, main = "FiberMargin")
par(old_par)
```

## Inspect pointwise diagnostics

```{r diagnostics}
head(data.frame(
  initial = sim$labels,
  refined = refined,
  candidate = attr(refined, "candidate"),
  margin_score = attr(refined, "margin_score"),
  required = attr(refined, "required"),
  repair_margin = attr(refined, "repair_margin"),
  atlas_dispersion = attr(refined, "atlas_dispersion"),
  changed = attr(refined, "changed")
))
```

`margin_score` is a local support contrast, and `repair_margin` is the margin above or
below the local acceptance threshold. Positive `repair_margin` means the candidate was accepted.
`atlas_dispersion` summarizes local route disagreement, and `changed` marks updated sites.

## Evaluate refinement quality

```{r evaluate}
evaluate_spatial_refinement(
  truth = sim$truth,
  initial = sim$labels,
  refined = refined,
  boundary = sim$boundary,
  regions = sim$area,
  sparse = sim$sparse,
  method = "FiberMargin"
)
```

Accuracy and ARI summarize global recovery.  
`macro_recall` and `worst_recall` expose class imbalance robustness.  
`boundary_accuracy` and `sparse_region_accuracy` inspect difficult spatial strata.  
`correction_recall` and `damage_rate` describe how much the algorithm repairs versus damages.

## Create benchmark geometries

```{r simulation-catalogue, eval=FALSE}
jagged <- simulate_spatial_domains(
  n = 50000L, pattern = "jagged_stripes", noise_type = "boundary",
  density_profile = "extreme", seed = 1L
)
gradient <- simulate_gradient_regions(
  n = 50000L, minority = 0.05, dimensions = 3, samples = 3L, seed = 2L
)
tubes <- simulate_complex_spatial_domains(
  n = 50000L, shape = "tubular_network", density_profile = "extreme", seed = 3L
)
volume <- simulate_volumetric_domains(
  n = 50000L, shape = "helical_channels", acquisition = "irregular_z", seed = 4L
)
```

```{r timed-benchmark, eval=FALSE}
benchmark_spatial_refiners(
  jagged,
  methods = list(
    FiberMargin = refine_spatial_labels,
    InitialOnly = function(xy, labels) labels
  )
)
```

## Use measured coordinates

For a Seurat object:

```{r seurat, eval=FALSE}
xy <- GetTissueCoordinates(object)[, c("x", "y")]
labels <- object$wsi_annotation_name
object$fibermargin <- refine_spatial_labels(xy, labels, workers = 2L)
```

For multiple specimens, concatenate arrays and pass a matching specimen identifier:

```{r multiple-specimens, eval=FALSE}
refined <- refine_spatial_labels(
  xy = coordinates,
  labels = initial_labels,
  samples = section_id,
  workers = 4L
)
```

```{r real-benchmark}
available_spatial_benchmarks()
dlpfc <- load_spatial_benchmark("dlpfc", scenario = 1L)
crc <- load_spatial_benchmark("crc", scenario = "CRC_random_25_r1")
c(spots = nrow(dlpfc$xy), dlpfc_sections = nlevels(dlpfc$samples))
c(spots = nrow(crc$xy), crc_classes = nlevels(crc$truth))
```

## Scope

FiberMargin is a selective refiner with noisy-label inputs only.
Coherent label swaps that keep the same boundary geometry are not observable from
coordinates alone, so accuracy should be interpreted together with `damage_rate`
and class-wise recall behavior.
