---
title: "FiberMargin workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FiberMargin workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 3)
library(fibermargin)
```

## Refine spatial labels

FiberMargin takes coordinates and initial categorical labels. A specimen
identifier is optional. The example contains four ordered areas whose reference
classes are A, B, B, and C, with minority labels mixed into every area.

```{r refine}
sim <- simulate_gradient_regions(
  n = 2400,
  minority = 0.25,
  samples = 2,
  density_profile = "strong",
  seed = 12
)

refined <- refine_spatial_labels(
  xy = sim$xy,
  labels = sim$labels,
  samples = sim$samples,
  workers = 1
)

c(
  input_accuracy = mean(sim$labels == sim$truth),
  refined_accuracy = mean(refined == sim$truth)
)
```

The method fixes its path atlas, central transport range, two-sided enclosure
field, class adjustment, local-gap protection, and chart-dispersion allowance
internally.
Specimens are processed independently, so translating, rotating, or relabeling one
specimen cannot affect another.

## Clean a categorical mask

A regular 2D mask or 3D voxel array can be supplied directly. FiberMargin uses only
array coordinates and categorical assignments; missing cells remain missing.

```{r clean-mask}
reference_mask <- matrix(rep(c("background", "region"), each = 800), nrow = 40)
noisy_mask <- corrupt_categorical_mask(
  reference_mask, mechanism = "impulse", rate = 0.15, seed = 19
)
cleaned_mask <- clean_categorical_mask(noisy_mask, workers = 1)

evaluate_mask_cleaning(reference_mask, noisy_mask, cleaned_mask)[, c(
  "initial_accuracy", "accuracy", "mean_iou", "mean_boundary_iou",
  "correction_recall", "damage_rate"
)]
```

The mask evaluator verifies the exact identity
`accuracy_gain = repaired_fraction - damaged_fraction`. This makes the correction
benefit and harm to initially correct cells visible in the same result.

## Visualize the correction

```{r spatial-plot, echo=FALSE, fig.cap="Reference regions, noisy input labels, and FiberMargin output for one specimen."}
rows <- sim$samples == levels(sim$samples)[1]
palette <- c(A = "#0072B2", B = "#E69F00", C = "#009E73")
old_par <- par(mfrow = c(1, 3), mar = c(1.5, 1.5, 2, 0.5))
plot(sim$xy[rows, ], col = palette[sim$truth[rows]], pch = 16, cex = 0.45,
     axes = FALSE, asp = 1, main = "Reference")
plot(sim$xy[rows, ], col = palette[sim$labels[rows]], pch = 16, cex = 0.45,
     axes = FALSE, asp = 1, main = "Noisy labels")
plot(sim$xy[rows, ], col = palette[refined[rows]], pch = 16, cex = 0.45,
     axes = FALSE, asp = 1, main = "FiberMargin")
par(old_par)
```

## Inspect decisions

The output is a factor with the input levels. Diagnostics are attached to each
location:

```{r diagnostics}
head(data.frame(
  initial = sim$labels,
  refined = refined,
  candidate = attr(refined, "candidate"),
  margin_score = attr(refined, "margin_score"),
  required = attr(refined, "required"),
  repair_margin = attr(refined, "repair_margin"),
  changed = attr(refined, "changed")
))
```

`margin_score` is an uncalibrated rival-versus-observed evidence contrast, not
a probability or expected loss. A nonnegative `repair_margin` accepts a
candidate that differs from the initial label.
`atlas_dispersion` measures disagreement among deterministic spatial charts. It
is not a sampling standard error or calibrated biological probability.

## Evaluate refinement

```{r evaluate}
evaluate_spatial_refinement(
  truth = sim$truth,
  initial = sim$labels,
  refined = refined,
  boundary = sim$boundary,
  regions = sim$area,
  sparse = sim$sparse,
  method = "FiberMargin"
)
```

Accuracy and adjusted Rand index summarize overall recovery. Macro and
worst-class recall expose imbalance; boundary and sparse-region accuracy score
difficult spatial strata; correction recall records repaired errors; and damage
records initially correct labels made wrong.

## Create benchmark geometries

All simulators return coordinates, noisy labels, reference labels, specimen IDs,
and evaluation masks.

```{r simulation-catalogue, eval=FALSE}
jagged <- simulate_spatial_domains(
  n = 50000, pattern = "jagged_stripes", noise_type = "boundary",
  density_profile = "extreme", seed = 1
)
gradient <- simulate_gradient_regions(
  n = 50000, minority = 0.05, dimensions = 3, samples = 3, seed = 2
)
tubes <- simulate_complex_spatial_domains(
  n = 50000, shape = "tubular_network", density_profile = "extreme", seed = 3
)
volume <- simulate_volumetric_domains(
  n = 50000, shape = "helical_channels", acquisition = "irregular_z", seed = 4
)
```

`benchmark_spatial_refiners()` times named methods on the same object. Its default
method is FiberMargin.

```{r timed-benchmark, eval=FALSE}
benchmark_spatial_refiners(jagged)
```

## Use measured coordinates

For a Seurat object, extract coordinates and pass the annotation vector directly.

```{r seurat, eval=FALSE}
library(Seurat)

xy <- GetTissueCoordinates(object)[, c("x", "y")]
labels <- object$wsi_annotation_name
object$fibermargin <- refine_spatial_labels(xy, labels)
```

For several sections, concatenate their coordinates and labels and pass a
matching section identifier:

```{r multiple-specimens, eval=FALSE}
refined <- refine_spatial_labels(
  xy = coordinates,
  labels = initial_labels,
  samples = section_id,
  workers = 4
)
```

The package bundles compact DLPFC and CRC Visium HD coordinate/annotation benchmarks.
CRC denotes colorectal cancer. The CRC derivative is CC BY 4.0 and excludes expression
counts and imagery.
The processed MERFISH annotation remains external because its redistribution terms
are not explicit.

```{r real-benchmark}
available_spatial_benchmarks()
dlpfc <- load_spatial_benchmark("dlpfc", scenario = 1)
c(spots = nrow(dlpfc$xy), sections = nlevels(dlpfc$samples))

crc <- load_spatial_benchmark("crc", scenario = "CRC_random_25_r1")
c(spots = nrow(crc$xy), classes = nlevels(crc$truth))
```

## Interpret the limit

Coordinates and one observed label cannot identify every error. A coherently
swapped region can be observationally indistinguishable from a genuine region.
FiberMargin is therefore evaluated as a selective refiner, not an annotation
oracle, and performance must be reported together with damage and weakest-class
behavior.
